\chapter{Properties of RACL}
\label{cha:properties_of_racl}

% Write about the basic ideas behind the proof of preservation and progress

We are now finally ready to state and prove the progress and preservation
properties of RACL. These are then used to prove
quasi-determinism. 


\section{Preservation \& Progress}%
\label{sec:preservation_and_progress}

The preservation theorem states that if we have a well-typed state that can
step to another state, this will also be well-typed.

\begin{theorem}[Preservation]
  \label{thm:preservation}
  Let $S, S'$ be states such that $\vdash S \tsep~\stateok$ and $S \Rrightarrow
  S'$. Then $\vdash S' \tsep \stateok$.
\end{theorem}

The proof of preservation in its full can be found in
Appendix~\ref{cha:proof_of_pnp}. The part concerning preservation of isolation
is probably the most interesting since properties like well-typed heap and
threads are standard. The preservation proof of isolation is based on the
fact that, by OCAP reachability, all $\ocap$ threads can only reach $\ocap$
class objects. Combined with global object separation, this means that an
$\ocap$ thread can never access the global object or create non-$\ocap$ objects
through a $\New{...}$-expression. This implies that the thread follows the OCAP model
as described in Section~\ref{sec:the_object_capability_model}. Thus isolated
threads will stay isolated.

Theorem~\ref{thm:progress} states that a well-typed program never gets stuck
unless we encounter an error or finish execution properly.
\begin{theorem}[Progress]
  \label{thm:progress}
  Let $S$ be a state such that $\vdash S \tsep \stateok$. Then either 
  \begin{enumerate}
    \item $\exists S'$ s.t. $S \Rrightarrow S'$, 
    \item $S = H, \emptyset$ for some heap $H$ s.t. $\noSpawn{(H)}$ or
    \item $S = \Error$.
  \end{enumerate}
\end{theorem}

The proof of progress is by contradiction. The full proof can also be found in
Appendix~\ref{cha:proof_of_pnp}.
% Describe:
% Theorem of determinism
% Lemmas leading up to the proof of this

\section{Quasi-Determinism}%
\label{sec:quasi_determinism}

Having proved soundness of the type system, the next thing to prove is
quasi-determinism. In order to state this more formally, we need some
definitions.

We first define a relation on states and show that it is an equivalence
relation.
\begin{definition} \label{def:eqrel}
  Let $\simeq$ be a binary relation on the set of states $\States$.
  We let $S \simeq S'$ if
  \begin{equation*}
    S = S' = \Error
  \end{equation*}
  or if
  \begin{equation}
    S = H, P \andalso S' = H', P'
  \end{equation}
  and there exists bijections $g \in \OIDs \hookrightarrow \OIDs, h \in \TIDs
  \hookrightarrow \TIDs$ such that
  \begin{equation}
    H' = \pi(H, g) \andalso P' = \rho(P, g, h)
  \end{equation}
  
  The functions $\pi$ and $\rho$ replace all object and thread identifiers
  occuring in $H$ or $P$ according to the bijections $g$ and $h$. They are
  explicitly defined in Definition~\ref{def:pirho},
  Appendix~\ref{cha:proof_of_qd}.
\end{definition}



\begin{proposition} \label{prop:eqrel}
  $\simeq$ is an equivalence relation.
\end{proposition}

A proof sketch of this can be found in Appendix~\ref{cha:proof_of_qd}.
Similarly we can prove the following. 
\begin{proposition} \label{prop:eqrel_stateok}
  For any $S, S' \in \States$ such that $S \simeq S'$
  \begin{equation}
    \vdash S \tsep \stateok \iff \vdash S' \tsep \stateok
  \end{equation}
\end{proposition}
A proof sketch can also be found in Appendix~\ref{cha:proof_of_qd}.
We can also make the following observation
\begin{claim}
  A transition between two states can be uniquely identified by:
  \begin{itemize}
    \item A start state $S$
    \item A base rule name $R$, e.g., $R = \ENew$ or $R = \ESpawn$. We say base rule
      because in many cases the step involves more than one rule in its
      derivation tree of a complete state step. However the use of these extra
      rules are implied by which rule is used as the "base" of the tree. For
      example, using $\ECall$ to reduce a thread implies use of $\EFSProp$ to
      advance the state, and using $\ENew$ to advance one thread frame implies
      use of $\EFProp$ and $\EFSProp$ in order to advance the state.
    \item A thread identifier $\iota \in \TIDs$. In the cases where we advance or terminate
      a thread, e.g., $\ENull$ or $\ECall$, this is the thread identifier of
      that thread, as in Definition~\ref{def:thread_sets}. In the case where we
      spawn a new thread, this is the unique thread identifier as in
      Definition~\ref{def:heap_obj}.
    \item A fresh object or thread identifier $\beta$ which could be of value:
      \begin{itemize}
        \item Object identifier $o' \in \OIDs$ in the cases where the rule
          references a fresh object identifier, i.e. $\ENew$ and $\ENewCell$.
        \item Thread id $\iota' \in \TIDs$ for the case where a fresh thread id is
          referenced, i.e. $\EWhen$.
        \item $\smiley$, a default value for all other cases.
      \end{itemize}
  \end{itemize}
\end{claim}

Because of this claim we can make the following definition

\begin{definition} \label{def:trans_id}
  We write a \emph{transition identifier} as $R^{\iota, \beta}$. The
  \emph{application} of $R^{\iota, \beta}$ to state $S$ (if possible) is the
  use of a rule specified by $R$, $\iota$ and $\beta$ as described above, to
  step from $S$ to some state $S'$. We write this as
  \begin{equation*}
    S \Rrightarrow^{R^{\iota, \beta}} S'
  \end{equation*}
\end{definition}

\begin{remark}
  The rule name $R$ in the above definition is actually redundant since this is
  completely decided by the state $S$ and the thread identifier $\iota$.
\end{remark}

To simplify notation we make the following definition.

\begin{definition}
  A \emph{transition sequence} $\bar{R}$ is a finite length sequence of
  transition identifiers,
  \begin{equation*}
    \bar{R} = R_1^{\iota_1, \beta_1}, R_2^{\iota_2, \beta_2}, \dots,
    R_n^{\iota_n, \beta_n}.
  \end{equation*}
  The application of this sequence to a state $S$ is the consecutive application
  of
  \begin{equation*}
    R_i^{\iota_i, \beta_i} \andalso i = 1, \dots, n
  \end{equation*}
  beginning with state $S$. I.e.
  \begin{equation*} 
    S \Rrightarrow^{R_1^{\iota_1, \beta_1}} S_1 \Rrightarrow^{R^{\iota_2,
    \beta_2}} S_2  \: \dots \: S_{n-1} \Rrightarrow^{R_n^{\iota_n, \beta_n}} S_n
  \end{equation*}
  We shorten this to
  \begin{equation*}
    S \Rrightarrow^{\bar{R}} S_n.
  \end{equation*}
  We call $n$ the length of $\bar{R}$.
\end{definition}

We state the following proposition without proof, just for reference. This
follows from the observations made in the claim above.
\begin{proposition} \label{prop:uniq_trans}
  Let $H, P \Rrightarrow H', P'$. Then there is a unique transition identifier
  $R^{\alpha, \beta}$ such that $H, P \Rrightarrow^{R^{\alpha, \beta}} H', P'$.
\end{proposition}

We are now ready to state the last main theorem.

\begin{theorem} \label{thm:qd}
  Let $S, S', T, T'$ be well-typed states not equal to $\Error$ such that
  \begin{equation*}
    \begin{gathered}
      S \simeq S' \\
      S \Rrightarrow^{\bar{R}} T \andalso S' \Rrightarrow^{\bar{R'}} T',
    \end{gathered}
  \end{equation*}
  Let $n$ and $n'$ be the lengths of $\bar{R}$ and $\bar{R'}$ respectively.
  Furthermore we assume that neither $T$ or $T'$ can make a step. Then
  \begin{equation*}
    n = n' \quad \text{and} \quad T \simeq T'.
  \end{equation*}
\end{theorem}

The proof of Theorem~\ref{thm:qd} is based on induction on the length of the two
transition sequences. It can be seen as incrementally transforming $\bar{R'}$ to
$\bar{R}$ by moving transition identifiers to positions specified by $\bar{R}$.
Each intermediate sequence is shown to halt in the same state. The full proof
can be found in Appendix~\ref{cha:proof_of_qd}.

Finally, quasi-determinism follows from Theorem~\ref{thm:qd} and
proposition~\ref{prop:uniq_trans}:
\begin{corollary}[Quasi-Determinism]
  Let $S, S', T, T'$ be well-typed states not equal to $\Error$ such that
  \begin{equation*}
    \begin{gathered}
      S \simeq S' \\
      S \Rrightarrow^* T \andalso S' \Rrightarrow^* T',
    \end{gathered}
  \end{equation*}
  and neither $T$ or $T'$ can make a step.  Then
  \begin{equation*}
    T \simeq T'.
  \end{equation*}
\end{corollary}

\section{Summary}%
\label{sec:summary}

We have proved soundness for the formal model RACL. This means that, given that
a program is well typed under our type system, it will never reach an
error state unless it encounters a null-pointer exception. Furthermore, the
program will never halt unless it has reached an error state, or no threads are
running and no threads are waiting to spawn.

Furthermore we have proved that RACL posesses the property of quasi-determinism.
This means that all non-erroneous halting states of a well-typed program are
equivalent.

%
%\begin{remark}
%  You should be able to prove that our system posesses a property similar to the
%  confluence property proved by \textcite{confluence}. This should not be hard
%  using the 
%\end{remark}
%
