\documentclass{kththesis}

\usepackage{csquotes} % Recommended by biblatex
\usepackage{biblatex}
\addbibresource{references.bib} % The file containing our references, in BibTeX format

\usepackage{mathtools, graphicx, hyperref, amssymb, algorithm2e}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{bcprules}
\usepackage{multicol}
\usepackage{tikz}
\usepackage{mathrsfs}
\usepackage{amsthm}
\usepackage{scrextend}
\usepackage{stmaryrd}
\usepackage{wasysym}
\usepackage{float}
\usepackage[labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{enumitem}

\usepackage{inconsolata}

\setlist[description]{listparindent=\parindent}

% tikz packages
\usetikzlibrary{arrows.meta,decorations.pathreplacing}

% Configuration of font and stuff for listings
\input{listings_settings.tex}

% Definitions of commands
\input{definitions.tex}

% Declarations of theorems and alike
\input{thmdecls.tex}

\title{Concurrent Determinism Using Lattices and the Object Capability Model}
\alttitle{Determinism i parallelliserade program med hjälp av gitterstrukturer och objektsförmågor}
\author{Ellen Arvidsson}
\email{magarv@kth.se}
\supervisor{Philipp Haller}
\examiner{Mads Dam}
\programme{Master in Computer Science}
\school{School of Electrical Engineering and Computer Science}
\date{\today}

\widowpenalty10000
\clubpenalty10000

\begin{document}

% Frontmatter includes the titlepage, abstracts and table-of-contents
\frontmatter

\titlepage

\begin{abstract}
  Parallelization is an important part of modern data systems. However, the
  non-determinism of thread scheduling introduces the difficult problem of
  considering all different execution orders when constructing an algorithm.
  Therefore deterministic-by-design concurrent systems are attractive.  A new
  approach called LVars consists of using data which is part of a lattice, with
  a predefined join operation. Updates to shared data are carried out using the
  join operation and thus the updates commute. Together with limiting the reads
  of shared data, this guarantees a deterministic result. The Reactive Async
  framework follows a similar approach but has several aspects which can cause a
  non-deterministic result. The goal with this thesis is to explore how we can
  ammend Reactive Async in order to guarantee a deterministic result. First an
  exploration into the subtleties of lattice based data combined with
  concurrency is made.  Then a formal model based on a simple object-oriented
  language is constructed.  The constructed small-step operational semantics and
  type system are shown to guarantee a form of determinism. This shows that
  LVars-similar system can be implemented in an object-oriented setting.
  Furthermore the work can act as a basis for future revisions of Reactive Async
  and similar frameworks.
\end{abstract}


\begin{otherlanguage}{swedish}
  \begin{abstract}
    Parallellisering är en viktig del i moderna datasystem. Flertrådade
    applikationer innebär dock en svårighet i och med att programmerare måste ta
    alla exekveringsordningar i beaktning. Därför är beräkningsmodeller vars
    resultat är garanterat deterministiskt en attraktiv utväg. En ny modell,
    kallad LVars, använder gitterstrukturer tillsammans med en
    supremum-operation för att garantera att uppdateringar av delad data
    kommuterar. Detta tillsammans med begränsningar av läsning av datan
    garanterar ett deterministiskt resultat. Reactive Async är ett
    programmeringsramverk som följer en liknande strategi. Det finns dock flera
    delar i dess konstruktion som i en oförsiktig programmerares händer kan
    orsaka att ett programs resultat blir icke-deterministiskt. Målet med detta
    examensarbete är att utforska vilka modifikationer som skulle kunna göras av
    Reactive Async för att garantera determinism. Först görs en undersökning av de mer
    svårförståeliga delarna i kombinationen av gitterbaserad data med
    flertrådad exekvering. Sedan kostrueras en formell beräkningsmodell baserad på
    ett enkelt objekt-orienterat språk. Konstruktionens småstegade operationella
    semantik tillsammans med dess typsystem visas kunna garantera en form av
    determinism. Detta visar att ett system liknande LVars kan implementeras i
    ett objekt-orienterat språk. Därmed skulle detta arbete kunna ligga till
    grund för framtida versioner av Reactive Async.
  \end{abstract}
\end{otherlanguage}


\tableofcontents


% Mainmatter is where the actual contents of the thesis goes
\mainmatter


%We use the \emph{biblatex} package to handle our references.  We therefore use the
%command \texttt{parencite} to get a reference in parenthesis, like this
%\parencite{heisenberg2015}.  It is also possible to include the author
%as part of the sentence using \texttt{textcite}, like talking about
%the work of \textcite{einstein2016}.

\include{./tex/chapter1}

\include{./tex/chapter2}

\include{./tex/chapter3}

\include{./tex/chapter4}

\include{./tex/chapter5}

\include{./tex/chapter6}

\include{./tex/chapter7}

\printbibliography[heading=bibintoc] % Print the bibliography (and make it appear in the table of contents)

\appendix

\include{./tex/appendixA}

\include{./tex/appendixB}

\end{document}
